---
title: "Acquisition and Cleaning of 23andMe Genomic Data"
author: "Philip Chase"
date: "3/29/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r read_data, echo=FALSE, results="hide", message=FALSE}
setwd('/Users/pbchase/classes/gms6804-group3-project')

df <- read.csv('combined_conversion_log.csv', stringsAsFactors=FALSE)
library(tidyr)
library(dplyr)
wide_df <- df %>% spread(attribute, value)
str(wide_df)
summary(wide_df)

# aggregate similar magic number strings to reduce number of factors
combine_magic_numbers <- function(x) {
  x <- gsub("^(ascii text).*", "\\1", x, ignore.case=TRUE)
  x <- gsub("^(ISO-8859 text).*", "ASCII text", x, ignore.case=TRUE)
  x <- gsub("^(Variant Call Format).*", "\\1", x, ignore.case=TRUE)
  x <- gsub("^(Composite Document File V2).*", "\\1", x, ignore.case=TRUE)
  x <- gsub("^(jpeg).*", "\\1", x, ignore.case=TRUE)
  x <- gsub("^(html).*", "\\1", x, ignore.case=TRUE)
  x <- gsub("^(SAMtools BAI).*", "\\1", x, ignore.case=TRUE)
  x <- gsub("^(bzip2).*", "\\1", x, ignore.case=TRUE)
  x <- gsub("^(gzip).*", "\\1", x, ignore.case=TRUE)
  x <- gsub("^(pdf).*", "\\1", x, ignore.case=TRUE)
  x <- gsub("^(SQLite 3.x database).*", "\\1", x, ignore.case=TRUE)
  x <- gsub("^(zip).*", "\\1", x, ignore.case=TRUE)
}

wide_df$combined_file_type <- combine_magic_numbers(wide_df$actual_file_type_from_magic_number)
wide_df$combined_file_type_of_member <- combine_magic_numbers(wide_df$member_type_from_magic_number)

# fix column classes
wide_df$archive_member_count <- as.integer(wide_df$archive_member_count)
wide_df <- wide_df %>% mutate_if(is.character,as.factor)
wide_df$subject <- as.character(wide_df$subject)
wide_df$member_name <- as.character(wide_df$member_name)
```

## Overview

Processed log data looks like 

```{r summarize_wide_df}
summary(wide_df)
```

I'm working on the outline to this document at https://docs.google.com/document/d/1Y2vzSo_cbuu22CYgn919ksKFWHEGim1O42pc43W6yYo/edit#
I'm using this to guide what stats and visuals I provide form the log data.

## Bar charts to make

* combined_file_type - aggregated file types of actual downloads
* combined_file_type_of_member - aggregated file types after extracting good content from archives
* text_file_subtype - categorization of the text files after extracting good content from archives
* archive_content - categorization of archive content

## histograms to make

* archive_member_count - could be very boring, but we should generate it to find out

## subject identifiers in file names of .zip files

We need a way to characterize the file naming patterns in the zip fiels for the subjects described by `archive_content == "name_is_genome_mumble_txt" | archive_content == "other_name"`. This could get into a discussion about identifiers and the risk of incidental exposure of identifiers in self-reported data.  

We will also need to review the data collection policies of the study to see if they were aware of this issue.

Here are the fields of interest:

```{r filenames_in_zip_files}
wide_df %>% filter(archive_content == "name_is_genome_mumble_txt" | archive_content == "other_name") %>% select(member_name)
```

