---
title: "Subject identifiers in file names inside .zip files"
author: "Philip Chase"
date: "4/13/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Note: this work is incomplete.  I considered reporting on the number of study participants who identified themselves in the uploaded 23andMe data.  After reviewing the informed consent and the study's educational it the study subjects were well-informed about the risks of self-identification. At that point I abandoned this analysis. 

## subject identifiers in file names inside .zip files

We need a way to characterize the file naming patterns in the zip fiels for the subjects described by `archive_content == "name_is_genome_mumble_txt" | archive_content == "other_name"`. This could get into a discussion about identifiers and the risk of incidental exposure of identifiers in self-reported data.  

We will also need to review the data collection policies of the study to see if they were aware of this issue.

Here are the fields of interest with the "genome*" naming pattern:

```{r filename_in_zip_file_is_genome_mumble}
wide_df %>% filter(archive_content == "name_is_genome_mumble_txt") %>% select(member_name) %>% tail()
```

To summarize the identifiers issue:

extract the name portion from the zip file member name



```{r }
# extract the name portion from the zip file member name
member_names <- wide_df %>% filter(archive_content == "name_is_genome_mumble_txt") %>% select(member_name) %>% head()

# remove the prefix and suffix form the name
trimmed_names <- gsub("_?_full_[0-9]+.txt", "", gsub("genome_", "", member_names$member_name, ignore.case=TRUE), ignore.case=TRUE)

# split the full names into name components
split_names <- strsplit(tolower(trimmed_names), "_")

# build a vector of surnames, given_names, initials, and suffixes
given_names <- c("marty", "Jen", "Emma")
surnames <- c("williams", "Anderson", "southwell")
suffixes <- c("Jr", "Sr", "I", "II", "III", "IV")
census_names <- tolower(c(given_names, surnames, suffixes, letters))

diffed_names <- lapply(split_names,  function(x) setdiff(x, y=census_names))

qty_of_unmatched_name_components <- sapply(diffed_names, length)

x<- c("genome_Juhan_Sonin_Full_20110524082809.txt", "genome_Andrew_Beeler_Full_20160320135452.txt")
```

Complete US census bureau data for given names through 2015 is available at https://www.ssa.gov/oact/babynames/names.zip

The top 1000 surnames from the 2010 census are available at https://www2.census.gov/topics/genealogy/2010surnames/Names_2010Census_Top1000.xlsx

```{r get_census_data}
# Download data file
given_name_scratch = '/Users/pbchase/classes/gms6804-group3-project/temp'
setwd(given_name_scratch)
dataURL <- "https://www.ssa.gov/oact/babynames/names.zip"
dataFile <- "names.zip"

if (!file.exists(dataFile)){
    print("Downloading data file")
    download.file(dataURL, dataFile, "curl", quiet=TRUE)
    # Extract the contents of the compressed file
    unzip(dataFile)
}

library(data.table)  
files <- list.files(path = given_name_scratch,pattern = ".txt")
temp <- lapply(files, fread, sep=",")
given_names_raw <- rbindlist(temp)
given_names <- tolower(unique(given_names_raw$V1))

# get surnames
given_name_scratch = '/Users/pbchase/classes/gms6804-group3-project/temp'
setwd(given_name_scratch)
dataURL <- "https://www2.census.gov/topics/genealogy/2010surnames/Names_2010Census_Top1000.xlsx"
dataFile <- "Names_2010Census_Top1000.xlsx"

if (!file.exists(dataFile)){
    print("Downloading data file")
    download.file(dataURL, dataFile, "curl", quiet=TRUE)
}

# library(xlsx)

# surnames_raw <- read.xlsx(dataFile, sheetName="2010_Top1000", start_row=4, start_column=1)

```


Here are the fields of interest with other naming patterns:

```{r filename_in_zip_file_is_other}
wide_df %>% filter(archive_content == "other_name") %>% select(member_name)
```

